The number of lattices of Voronoi's first kind

### 2014-06-04 findsuperbase/findsuperbase.tex
interestly the mimo matrices don't ever appear to be of first kind!

The plot shows the estimated (after 500 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4$.  Three different methods for generating the $n \times n$ basis matrix $B$ are considered: 
\begin{enumerate}
\item entries drawn independently and uniformly in the interval $[0,1]$, 
\item entries drawn independently and Gaussian with zero mean and variance one, and 
\item the $i,j$th entry satisfies
\[
B_{i,j} = \begin{cases}
R_{i,j} & 1 \leq i,j \leq h \\
-C_{i-h,j} & h < i \leq n, 1 \leq j \leq h \\
C_{i,j-h} & h < j \leq n, 1 \leq i \leq h \\
R_{i-h,j-h} & h < i,j \leq n,
\end{cases}
\]
where $h = \ceil{n/2}$ is the smallest integer greater than or equal to $n/2$ and $R_{i,j}$ and $C_{i,j}$ for $i,j = 1,\dots,h$ are independently Gaussian distributed with zero mean and variance one. 
\end{enumerate}
When $n$ is even, the third case corresponds with the model commonly used for MIMO channels.  In this case $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.

As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.  When $n=4$ roughtly 25\% of the lattices generated were of Voronoi's first kind. This was the case for all 3 basis generation methods.  These simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero (in some sense!) in dimension 4.

@@

\begin{figure}[tp]
\centering 
  \begin{tikzpicture}
    \begin{axis}[font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
      \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
      \addplot[mark=*,color=blue,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/gaussian};
      \addplot[mark=x,color=green,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
\legend{Uniform, Gaussian}
    \end{axis}
  \end{tikzpicture}
\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
\end{figure}

>>>>

The plot shows the estimated (after 500 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4$.  Three different methods for generating the $n \times n$ basis matrix $B$ are considered: 
\begin{enumerate}
\item entries drawn independently and uniformly in the interval $[0,1]$, 
\item entries drawn independently and Gaussian with zero mean and variance one, and 
\item the $i,j$th entry satisfies
\[
B_{i,j} = \begin{cases}
R_{i,j} & 1 \leq i,j \leq h \\
-C_{i-h,j} & h < i \leq n, 1 \leq j \leq h \\
C_{i,j-h} & h < j \leq n, 1 \leq i \leq h \\
R_{i-h,j-h} & h < i,j \leq n,
\end{cases}
\]
where $h = \ceil{n/2}$ is the smallest integer greater than or equal to $n/2$ and $R_{i,j}$ and $C_{i,j}$ for $i,j = 1,\dots,h$ are independently Gaussian distributed with zero mean and variance one. 
\end{enumerate}
When $n$ is even, the third case corresponds with the model commonly used for MIMO channels, that is, the basis matrix is of the form
\[
B = \left(\begin{array}{cc}
R & -C \\
C & R
\end{array}\right)
\]
where $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.

As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.  When $n=4$ roughtly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 500 bases generated by method 3 were of first kind.  I find this very interesting! The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattices bases.  This might suggest that lattices of first kind will (atleast not directly) be useful for MIMO.

@@

\begin{figure}[tp]
\centering 
  \begin{tikzpicture}
    \begin{axis}[font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
      \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
      \addplot[mark=square,color=blue,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/gaussian};
      \addplot[mark=x,color=green,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
\legend{Uniform, Gaussian, MIMO}
    \end{axis}
  \end{tikzpicture}
\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
\end{figure}
